Spatial Reasoning and Statistical Graphics

### 2014-10-20 Paper/Draft1/TestingVisualAptitude.Rnw
Tweaks to paragraph structure.

Table \ref{tab:scorecomparison} shows mean scores and standard deviation for ISU students and other populations. Values have been adjusted to accommodate differences in test procedures and sub-population structure; for instance,  some data is reported for a single part of a two-part test, or results are reported for each gender separately (Adjustment procedure is described in more detail in Appendix \ref{app:ScoreAdj}). Once these adjustments have been completed, it is evident that Iowa State undergraduates scored at about the same level as other similar demographics. In fact, both means and standard deviations of ISU students' scores are similar to the comparison groups, which were chosen from available demographic groups based on population similarity.

Comparison populuation data was chosen in the following manner: if comparison data was available for 9th and 12th grade students, we have compared Iowa State students' scores with the 12th grade students, as they are closer in age to college students. When data was available from college students and Army enlistees, we have compared ISU students to other college students, as college students are more likely to have similar gender distribution to ISU students.

\comment{Could you include some more details on how to get from the Unscaled scores to the scaled versions? I'm assuming that for the Card Rotation, the two parts have equal size, explaining the doubled score, but why not double the standard deviation? $\text{Var}(aX) = a^2 \text{Var}(X)$  ok I got the first one - because it is $X_1 + X_2$ not $2 X_1$. still working on understanding the second one, though ... it would be worthwhile to include a bit more detail and maybe put that in an appendix to the paper. If you want to, you could instead extend the footnotes a bit for now. ... so just let me try this: Let's assume that $X_M$ and $X_F$ are the scores based on 294 men, and 323 women (are those midscore s of the intervals reported?). with corresponding $s_M$ and $s_F$ for the reported standard deviation of the full test. Then half of the test would have half of the scores with half of the standard deviation. we have a breakdown of 17 girls and 19 boys. $
X = (X_M/2 \cdot 294 + X_F/2 \cdot 323) /(294+323) = \Sexpr{round(120.0/2*294/617 + 114.9/2*323/617, 1)}
$. For the Variance of $X$ we then have:
$\text{Var}(X) = \text{Var}(X_{M_1} + X_{M_2}) \cdot (294/617)^2 + \text{Var}(X_{F_1} + X_{F_2}) \cdot (323/617)^2 = \Sexpr{round(sqrt(30^2/2 * (294/617)^2 + 27.8^2/2 * (323/617)^2),1)}^2$
}

>>>>

Table \ref{tab:scorecomparison} shows mean scores and standard deviation for ISU students and other populations. Values have been adjusted to accommodate differences in test procedures and sub-population structure; for instance,  some data is reported for a single part of a two-part test, or results are reported for each gender separately (adjustment procedure is described in more detail in Appendix \ref{app:ScoreAdj}). Once these adjustments have been completed, it is evident that Iowa State undergraduates scored at about the same level as other similar demographics. In fact, both means and standard deviations of ISU students' scores are similar to the comparison groups, which were chosen from available demographic groups based on population similarity.

Comparison populuation data was chosen to most closely match ISU undergraduate population demographics. Thus, if comparison data was available for 9th and 12th grade students, we have compared Iowa State students' scores with the 12th grade students, as they are closer in age to college students. When data was available from college students and Army enlistees, we have compared ISU students to other college students, as college students are more likely to have similar gender distribution to ISU students.

\comment{\textit{We may want to include a table of scaled scores here, so that we don't just bust right into lineup discussion without contextualizing everything properly.}}
\newtext{Applying the scaling protocol discussed in section \ref{sec:scaling}, we see that the ranges of scaled lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task.}

% 
% \comment{Could you include some more details on how to get from the Unscaled scores to the scaled versions? I'm assuming that for the Card Rotation, the two parts have equal size, explaining the doubled score, but why not double the standard deviation? $\text{Var}(aX) = a^2 \text{Var}(X)$  ok I got the first one - because it is $X_1 + X_2$ not $2 X_1$. still working on understanding the second one, though ... it would be worthwhile to include a bit more detail and maybe put that in an appendix to the paper. If you want to, you could instead extend the footnotes a bit for now. ... so just let me try this: Let's assume that $X_M$ and $X_F$ are the scores based on 294 men, and 323 women (are those midscore s of the intervals reported?). with corresponding $s_M$ and $s_F$ for the reported standard deviation of the full test. Then half of the test would have half of the scores with half of the standard deviation. we have a breakdown of 17 girls and 19 boys. $
% X = (X_M/2 \cdot 294 + X_F/2 \cdot 323) /(294+323) = \Sexpr{round(120.0/2*294/617 + 114.9/2*323/617, 1)}
% $. For the Variance of $X$ we then have:
% $\text{Var}(X) = \text{Var}(X_{M_1} + X_{M_2}) \cdot (294/617)^2 + \text{Var}(X_{F_1} + X_{F_2}) \cdot (323/617)^2 = \Sexpr{round(sqrt(30^2/2 * (294/617)^2 + 27.8^2/2 * (323/617)^2),1)}^2$
% }
